Multimodal Information Extraction of Supermarket Leaflets

### 2025-02-25 paper/main.tex
Add recent changes.
Ilyi: Final Version.

% This must be in the first 5 lines to tell arXiv to use pdfLaTeX, which is strongly recommended.
\pdfoutput=1
% In particular, the hyperref package requires pdfLaTeX in order to break URLs across lines.

@@

% Change "review" to "final" to generate the final (sometimes called camera-ready) version.
% Change to "preprint" to generate a non-anonymous version with page numbers.
\usepackage[preprint]{coling}

% Standard package includes
\usepackage{times}
\usepackage{latexsym}

% For proper rendering and hyphenation of words containing Latin characters (including in bib files)
\usepackage[T1]{fontenc}
% For Vietnamese characters
% \usepackage[T5]{fontenc}
% See https://www.latex-project.org/help/documentation/encguide.pdf for other character sets

% This assumes your files are encoded as UTF8
\usepackage[utf8]{inputenc}

% This is not strictly necessary, and may be commented out,
% but it will improve the layout of the manuscript,
% and will typically save some space.
\usepackage{microtype}

% This is also not strictly necessary, and may be commented out.
% However, it will improve the aesthetics of text in
% the typewriter font.
\usepackage{inconsolata}

%Including images in your LaTeX document requires adding
%additional package(s)
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{amssymb}
\usepackage{hyperref}

@@

\author{Xincheng Liao\textsuperscript{}, Junwen Duan\textsuperscript{}\thanks{\ \ Corresponding author. Email: \href{mailto:jwduan@csu.edu.cn}{jwduan@csu.edu.cn}}, Yixi Huang\textsuperscript{}, Jianxin Wang\textsuperscript{} \\
Hunan Provincial Key Lab on Bioinformatics, School of Computer Science and Engineering, \\
Central South University, Changsha, Hunan, China \\
\texttt{\{ostars, jwduan, yx.huang\}@csu.edu.cn, jxwang@mail.csu.edu.cn} \\
\href{https://github.com/OStars/RUIE}{https://github.com/OStars/RUIE}
}

\begin{document}
\maketitle
\begin{abstract}
TODO
\end{abstract}

\section{Introduction}
\subsection{Motivation}
\subsection{Problem Statement}
\subsection{Structure of Paper}

\section{Related Works}
\indent{\textbf{Deal Detection}.}
\indent{\textbf{Optical Character Recognition}.}
OCR-Model-Driven methods use OCR tools to acquire text
and bounding box information. Subsequently, they rely
on the models to integrate text, layout, and visual data.

\indent{\textbf{Information Extraction}.}

\section{Methodological Overview}

\section{Deal Detection}
    \subsection{Datasets}
    \subsection{Model}
    \subsection{Experiments}

\section{Optical Character Recognition}

\section{Information Extraction}

\section{Application}

\section{Conclusion}
    \subsection{Summary}
    \subsection{Future Work}
    ww

% Bibliography entries for the entire Anthology, followed by custom entries
%\bibliography{anthology,custom}
% Custom bibliography entries only
\bibliography{references}

>>>>

% !TEX root = main_ilyi.tex

% for pdfLaTeX
% \pdfoutput=1

@@

% Change "review" to "final" to generate the final (sometimes called camera-ready) version.
% Change to "preprint" to generate a non-anonymous version with page numbers.
\RequirePackage{lib}

% Standard package includes
\RequirePackage{times}
\RequirePackage{latexsym}

% For proper rendering and hyphenation of words containing Latin characters (including in bib files)
\RequirePackage[T1]{fontenc}

% This assumes your files are encoded as UTF8
\RequirePackage[utf8]{inputenc}

\RequirePackage{microtype}

\RequirePackage{inconsolata}

\RequirePackage{graphicx}
\RequirePackage{amsmath}
\RequirePackage{multirow}
\RequirePackage{booktabs}
\RequirePackage{amssymb}
\RequirePackage{hyperref}
\RequirePackage{subfig}

\RequirePackage{algorithm}
\RequirePackage{algpseudocode}
\RequirePackage{verbatim}

@@

\author{Ilyesse Hettenbach\textsuperscript{}, Gabriel Schurr\textsuperscript{} \\
HKA, Karlsruhe, Germany \\
\texttt{\{heil1012, scga1011\}@h-ka.de} \\
\href{https://github.com/ilyii/leaflets}{https://github.com/ilyii/leaflets}
}

\begin{document}
\maketitle
\begin{abstract}
Supermarket leaflets are a valuable source of information for consumers, offering insights into product deals, discounts, and promotions. However, extracting structured information from these leaflets is a challenging task due to their unstructured nature and layout variability. In this work, we present an intelligent system for extracting, structuring, and utilizing information from supermarket leaflets in a digital format. Our approach leverages advanced computer vision, Optical Character Recognition (OCR), and language models to detect and extract deals from these leaflets. We propose a modular pipeline for deal detection and information extraction, develop an interactive application for browsing and comparing deals, and introduce Leaflet-IE, a benchmark dataset for evaluating information extraction models. Our system aims to enhance the digital transformation of supermarket leaflets, providing consumers with a more efficient and data-driven shopping experience.
\end{abstract}

\section{Introduction}

\subsection{Motivation}
When one thinks of a paper-heavy country, Germany is likely to come to mind. The country's affinity for printed materials is deeply ingrained in its culture, with activities such as browsing through supermarket leaflets serving as a familiar ritual for many individuals. These leaflets have long been an essential medium for consumers to plan their purchases, discover special offers, and compare prices.
The rapid advancement of digital technology is transforming consumer behavior. While some individuals still find browsing through physical leaflets a nostalgic and even relaxing experience, younger generations are increasingly turning to digital alternatives for their shopping needs. The modern shopper expects convenience, efficiency, and instant access to information, yet the digital transformation of supermarket leaflets in Germany lags significantly behind, even despite the existence of applications that centralize digital versions of these leaflets. The majority of these applications' digital leaflets are plain documents, failing to leverage the full potential of interactivity, searchability, and intelligent recommendation systems.

Furthermore, while price comparison websites and applications have become commonplace for electronics, fashion, and travel, the grocery sector remains an overlooked frontier. Consumers are often left navigating multiple supermarket websites or manually cross-referencing prices to find the best deals—a tedious and inefficient process. A comprehensive digital ecosystem that integrates grocery price comparisons, personalized discounts, and AI-driven shopping assistants could bridge this gap and act as a workaround, offering an enhanced, data-driven shopping experience and a tool that allows consumers to create a dynamic shopping list, track price trends over time, and receive real-time updates on the best deals.

\subsection{Problem Statement}
The primary objective of this research is to develop an intelligent system for extracting, structuring, and utilizing information from supermarket leaflets in a digital format. Given a set of supermarket leaflets $\mathbf{L} = \{L_1, L_2, ..., L_n\}$, where each $L_i$ contains a set ofproduct deals $\{p_j\}_{j=1}^{m} \in P_{L_i}$, the goal is to extract and normalize these attributes to enable efficient querying and comparison across multiple supermarket chains.

Each product deal $p_j$ is represented by a set of attributes $\{y_1, y_2, ..., y_k\} \subset p_j$, where each attribute $y_k$ represents a specific entity. Common attributes include the name and brand of the product, the original and deal prices, the unit (e.g., weight, volume), and the product image. The challenge lies in detecting and separating deals from the leaflet, extracting structured information from the detected deals, normalizing the extracted information to ensure consistency and comparability, and validating the extracted information to ensure accuracy.

To enable advanced applications such as the abovementioned, one has to tackle different challenges, such as:
\begin{enumerate}
    \item Detecting and separating deals $p$ from the leaflet $L$.
    \item Extracting structured information from the detected deals.
    \item Normalizing the extracted information to ensure consistency and comparability.
    \item Validating the extracted information to ensure accuracy.
\end{enumerate}

Therefore, a function $f: \mathcal{L} \to \mathcal{D}$ has to be designed, that maps a set of leaflets $\mathcal{L}$ to a set of structured deals $\mathcal{D}$, where each deal $d \in \mathcal{D}$ is represented by a set of attributes $\{y_1, y_2, ..., y_k\}$.

\subsection{Contributions}
To build this mapping, this work presents the conducted research and the practical result of a comprehensive approach to digitalizing and structuring supermarket leaflet information by leveraging advanced computer vision, Optical Character Recognition (OCR), and language models, depicted in \figref{fig:overview}. A \emph{modular pipeline} is proposed that encompasses deal detection, information extraction and normalization to transform raw leaflet data into structured, actionable insights.

\begin{figure*}
    \centering
    \includegraphics[width=0.8\linewidth]{figures/overview.png}
    \caption{Overview of the proposed approach to digitalizing and structuring supermarket leaflet information.}
    \label{fig:overview}
\end{figure*}

To enhance usability, an \emph{interactive application} is developed that allows consumers to browse, filter, and compare deals across multiple supermarket chains. It provides essential search, tracking, and comparison features for improved deal discovery.

To facilitate research and development in this domain, \emph{Leaflet-IE}, an end-to-end benchmark dataset for information extraction from supermarket leaflets, is introduced. This dataset stands as a valuable resource for evaluating and comparing different models and approaches in the field.

\subsection{Structure}
This work is structured to systematically address supermarket deal detection and information extraction, beginning with a scientific overview of related works in \secref{sec:related_works}. The nature of supermarket leaflet data is analyzed in \secref{sec:leaflets_data}, formalizing challenges such as layout variability, OCR noise, and multimodal fusion. Deal detection is covered in \secref{sec:deal_detection}, detailing an own dataset creation, instance segmentation model selection, and experimental results for identifying promotional content. Information extraction is extended in \secref{sec:information_extraction}, detailing an own benchmark dataset construction, popular architectural approach and their performance evaluation for extracting structured product and pricing information. Thereafter, the proposed interactive application is introduced in \secref{sec:application}, outlining its features, design, and user interface. The work concludes with a summary of contributions and future research directions in \secref{sec:conclusion}.

\section{Related Works}
\label{sec:related_works}

\indent{\textbf{Supermarket Leaflet Analysis}.}
The field of supermarket leaflet analysis has evolved to tackle key challenges such as product classification, promotion identification, and customer interaction. Recent studies have leveraged AI-driven techniques to enhance efficiency and accuracy in this domain. For instance, cite{ladwig2023} explore fine-grained product classification in leaflet advertisements, while \cite{arroyo2020} employ multi-label classification methods to analyze promotions using both textual and visual information. Additionally, Zhang et al. \cite{zhang2024} introduce PKU-GoodsAD, a dataset designed for unsupervised anomaly detection and segmentation of supermarket goods. Complementing these efforts,\cite{nandkumar2024} propose a multi-level Large Language Model (LLM) conversational interface to enhance supermarket robot interactions by handling diverse customer intents.

\indent{\textbf{Deal Detection}.}
There are no direct approaches for detecting deals, including product image, price, name and more, in supermarket leaflets. However, the detection of specific elements in documents, or more generally, the detection of objects in images, is a widely researched area. When it comes to just the actual detection of these elements without understanding the semantic meaning, models such as YOLO or Faster R-CNN prove to be effective. These models are able to recognize objects in images and mark them with bounding boxes and further classify them. But because there is no direct research on the detection of deals in supermarket leaflets, the research has to be done in this area.

\indent{\textbf{Information Extraction}.}
Recent advancements in OCR, LLMs, and end-to-end multimodal architectures have significantly refined the landscape of document information extraction. These improvements have led to increased accuracy, efficiency, and versatility in processing structured and unstructured text from diverse document formats.

OCR technology has evolved from traditional rule-based approaches to sophisticated deep learning-driven systems. Open-source OCR frameworks such as Tesseract \cite{tesseract}, PaddleOCR \cite{ppocrv2}, EasyOCR \cite{easyocr}, and DocTR \cite{doctr} integrate state-of-the-art methodologies to enhance both text detection \cite{liao2019, liao2022} and recognition \cite{shi2015, li2022}. These systems employ deep neural networks to improve accuracy in complex document layouts, including distorted and low-quality images. Preprocessing techniques play a crucial role in optimizing OCR performance. Methods such as rectification \cite{shi2019}, deskewing \cite{pham2022}, and denoising \cite{zhao2018} significantly enhance text clarity by mitigating distortions, reducing noise, and improving alignment. These preprocessing steps are especially useful in real-world scenarios where scanned or photographed documents contain various inconsistencies. LLMs have emerged as powerful tools for refining OCR outputs through context-aware reasoning and semantic corrections. Models like GPT-3 \cite{brown2020} and LLaMA \cite{touvron2023} demonstrate their performance in natural language understanding, enabling improved entity recognition and post-processing of extracted text \cite{yenduri2023}. These models enhance the reliability of OCR by resolving ambiguities, correcting misinterpretations, and providing structured outputs in diverse domains such as finance, healthcare, and legal documentation.

Transformer-based architectures \cite{vaswani2017, dosovitskiy2021} have also revolutionized document understanding by integrating vision and language processing into unified frameworks. These models leverage attention mechanisms to jointly process textual and visual cues, achieving state-of-the-art performance in document analysis \cite{kim2022,appalaraju2021}. These models are particularly effective in handling complex layouts, handwriting recognition, and tabular data extraction. Furthermore, multimodal approaches \cite{peng2022, li2021} introduce hybrid learning techniques that enhance information retrieval from scanned documents and digital forms.

The integration of large vision-language models (LVLMs) has further expanded the capabilities of document intelligence since they utilize large-scale multimodal training to perform sophisticated image-text reasoning \cite{touvron2023,qwen2025,dai2023}. These advancements enable robust document understanding, particularly in scenarios where textual information is closely intertwined with graphical elements. Recent contributions \cite{li2024, wei2024} emphasize the growing potential of LVLMs.

\section{The Nature of Supermarket Leaflet Data}  
\label{sec:leaflets_data}
Supermarket leaflets constitute a complex multimodal data domain characterized by heterogeneous layouts, ambiguous associations, and noisy signal extraction. As illustrated in \figref{fig:leaflet_data_samples}, these documents present unique computational challenges that demand joint reasoning over visual, textual, and contextual modalities. We formalize the key obstacles below.

Leaflets exhibit extreme variability in design elements that impede automated parsing. The spatial heterogeneity, thus the non-standardized arrangements of product images, prices, and promotions across retailers, is a primary concern. The layout of leaflets can vary significantly, with product images, prices, and promotional text appearing in diverse configurations. Additionally, typography challenges, such as superscripted decimals, struck-through original prices, and font inconsistencies between deals and regular text, further complicate the extraction process. Overlapping elements, where text is occluded by product images or graphical overlays, violate the assumptions of Optical Character Recognition (OCR) systems, which assume planar text.

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.8\linewidth]{figures/samples.png}
    \caption{Common German supermarket deals.}
    \label{fig:leaflet_data_samples}
\end{figure}

Inherently, these leaflets contain both textual and visual information, which must be jointly processed to extract meaningful insights. The textual elements include product names, brands, prices, and units, while the visual components consist of product images, logos, and promotional graphics. The challenge lies in aligning these modalities to extract structured information accurately. Furthermore, the multimodal nature of the data introduces semantic ambiguities, where textual and visual elements maintain non-strict correspondences. For instance, a product image may correspond to multiple text entries, or price interpretation may require visual grounding.

Due to these characteristics, leaflets are prone to linguistic ambiguities, conversion errors and cultural-specific semantics. The German language, with its compound nouns, specific punctuation conventions, and diverse word forms, poses unique challenges for text extraction. OCR errors, such as glyph confusion and tokenization failures, are common due to font artifacts and compound nouns. Multilingual mix, where English promotion terms appear alongside German body text, further complicates the parsing process. Price disambiguation, distinguishing unit prices from total prices without weight references, is a critical challenge. Additionally, more complex offer strategies, such as buy-one-get-one-free, coupon-based discounts, and permanent price reductions, require nuanced interpretation beyond simple text extraction.

Additionally, supermarket leaflets are commonly scheduled for a weekly release, with each leaflet containing a set of deals that are valid for a specific period. However, not every supermarket chain follows this pattern, and some may release leaflets bi-weekly, monthly or even irregularly. This temporal grounding problem requires associating each deal with an expiry time, which is often implicit in the leaflet text. For instance, a deal may be valid only in specific store branches or for a limited time, indicated by small print codes or context clues.

This multifaceted nature demands architectures that jointly optimize for visual-textual alignment, error-resilient parsing, and contextual plausibility.

\section{Deal Detection}
\label{sec:deal_detection}
Detecting deals within supermarket leaflets is a crucial step in the process of extracting structured  data. The ultimate end-to-end goal of deal detection is to accurately identify and separate individual deals from digital supermarket leaflets, enabling further analysis and integration into a structured database for an application. This step is fundamental for creating a comprehensive, queryable system that allows users to compare deals across different retailers efficiently. Directly detecting deals within leaflets is necessary because supermarket deals are not standardized, and each leaflet can contain multiple overlapping deals, diverse layouts, and varying fonts, making simple text-based extraction approaches insufficient.

\subsection{Dataset}
An approach to obtaining deal data would be direct web scraping from supermarket websites, as these days they present their deals digitally in a structured way. However, this method is highly error-prone and difficult to generalize due to the distinct and frequently changing layouts of different supermarket websites. Additionally, many retailers do not provide direct, easily scrappable structured deal data. Instead, an alternative source has been identified. Websites that aggregate and upload full promotional leaflets from various German supermarkets on a weekly basis. These leaflets provide a more standardized input format and can be systematically downloaded using an automated script.

The collected dataset consists of approximately 13,000 images from a diverse range of supermarket chains, including well-known retailers such as Penny, Aldi, and Kaufland, as well as less conventional ones like Metro. The primary focus remains on supermarkets that sell groceries and drugstore products, as these are the most relevant for the intended applications. Each leaflet page contains multiple distinct offers, which must be individually detected. This is achieved through object detection techniques, which can either utilize bounding boxes or instance segmentation. While bounding box detection provides a straightforward way to localize deals, it often struggles with overlapping offers, leading to incorrect classifications. Instance segmentation is therefore the preferred approach, as it allows for more precise extraction of deal regions without interference from neighboring offers. Deals in the dataset are categorized into four main classes:

\begin{itemize}
    \item Standard Deal: A regular product discount.
    \item Product Category Deal: A discount on an entire product category (e.g., "20\% off all canned goods").
    \item Coupon Deal: Special offers requiring the use of coupons (e.g., Payback).
    \item Miscellaneous: Deals outside the primary categories, such as travel promotions.
\end{itemize}

\begin{figure*}[h!]
    \centering
    \includegraphics[width=0.7\linewidth]{figures/deal_detection/dataset_count.png}
    \caption{Distribution of labeled leaflet pages and deals in training and validation sets.}
    \label{fig:ddetect_dataset_count}
\end{figure*}

The primary focus is on standard deals, as they are the most common and directly useful for consumers. No publicly available dataset fully aligns with this task as it involves both instance segmentation and deal categorization, specifically in the context of German supermarkets. So, a subset of the collected pages was manually annotated. The result is a dataset of 672 annotated leaflet pages, with 4,675 individual deals in training and 75 pages with 523 deals in the validation set as shown in \figref{fig:ddetect_dataset_count}. In cases where no overlapping elements were present, bounding boxes were used. For overlapping offers, instance segmentation masks were created. Depending on the training requirements, annotations can be converted between bounding boxes and segmentation masks, although segmentation masks provide a more granular representation.

\begin{figure*}[h!]
    \centering
    \includegraphics[width=0.6\linewidth]{figures/deal_detection/bbox_page.png}
    \caption{Bounding box detection results on a leaflet page resulting in overlapping bounding boxes.}
    \label{fig:ddetect_bbox_page}
\end{figure*}

\begin{figure*}[h!]
    \centering
    \includegraphics[width=0.6\linewidth]{figures/deal_detection/mask_page.png}
    \caption{Instance segmentation results on a leaflet page showing precise deal separation.}
    \label{fig:ddetect_mask_page}
\end{figure*}

\subsection{Model}
Object detection and segmentation tasks require identifying objects within an image while determining their locations. The simplest form of image analysis, image classification, assigns a single label to an entire image, providing no information about object positions. However, multiple deal detection in supermarket leaflets demands the identification of multiple deals, whose number is unknown beforehand, along with their precise locations to extract them later.

Object detection models can be broadly categorized into multi-stage (multi-shot) detectors and single-stage (single-shot) detectors. Multi-stage detectors, such as the R-CNN family (R-CNN, Fast R-CNN, Faster R-CNN, Mask R-CNN), follow a sequential process. First, they generate a set of region proposals, indicating potential object locations. These proposals are then classified in a second stage. While this method achieves high accuracy, it introduces significant computational overhead. Earlier versions, such as R-CNN, even relied on separate, non-learnable algorithms for proposal generation. Faster R-CNN improved efficiency by integrating the Region Proposal Network (RPN) directly into the model, eliminating the need for external proposal algorithms. However, even this approach remains computationally demanding. In contrast, single-stage detectors, such as YOLO (You Only Look Once), perform detection in a single forward pass, predicting object locations and classifications simultaneously. This significantly reduces inference time, making these models well-suited for real-time applications. YOLO, in particular, has established itself as one of the most widely used architectures due to its speed and competitive accuracy. Given the requirement of processing thousands of leaflet pages efficiently, a single-stage detector was chosen for deal detection, with a particular focus on YOLO-based architectures.

The YOLO model follows a grid-based detection approach, where the input image is divided into an SxS grid. Each grid cell is processed for predicting objects whose centers fall within it. Instead of relying on separate region proposal mechanisms, YOLO processes the image in a single forward pass, directly predicting bounding box coordinates, confidence scores, and class probabilities per grid cell. The bounding boxes are parameterized by their center coordinates, width, and height, while the confidence scores indicate the model’s certainty that a given bounding box contains an object. The class probabilities further specify which category the detected object belongs to. Traditional object detection frameworks like YOLO produce bounding boxes that encapsulate detected objects in rectangular shapes. However, supermarket leaflets contain deals that frequently overlap or feature irregular layouts, making bounding box-based detection less effective. Instance segmentation extends YOLO by incorporating an additional segmentation head that predicts pixel-level masks for each detected object. This approach enables the model to distinguish individual deals more precisely, reducing the risk of grouping multiple offers within a single bounding box. The segmentation mechanism operates by first detecting objects using the standard YOLO pipeline. Once a bounding box is identified, the model refines its prediction by generating a binary mask that determines which pixels within the bounding box belong to the detected object. The segmentation head shares feature representations with the object detection components, ensuring efficient training and inference. This method enhances the model’s ability to isolate individual deals even in cluttered layouts, making it particularly suited for supermarket leaflets where offers frequently overlap or appear with complex backgrounds.

\subsection{Experiments}
Evaluating the performance of instance segmentation models requires an analysis of how well detected objects align with ground-truth annotations. The primary metric used in this assessment is \emph{Mean Average Precision (mAP)}, which quantifies detection accuracy at different levels of overlap between predicted and actual objects. The core component of mAP is Intersection over Union (IoU). IoU is defined as the ratio of the intersection area between the predicted and ground-truth segmentation masks to their union. A higher IoU indicates a better match between the prediction and the actual object, while a lower IoU suggests that the detected region is misaligned or includes excessive background. In the case of bounding boxes, IoU is calculated based on the overlap of rectangular regions, whereas for instance segmentation, it is computed on a pixel level, capturing the true shape of the detected objects. To evaluate model performance, different IoU thresholds are applied. A commonly used metric is mAP@50, which considers a detection correct if the IoU between the predicted and ground-truth mask exceeds 50\%. This threshold provides insight into how well a model identifies objects in general but does not strictly enforce precision in localization. A more comprehensive metric is mAP@50-95, which averages precision across multiple IoU thresholds, ranging from 50\% to 95\% in increments of 5\%. This approach penalizes models that fail to produce accurate segmentations and is particularly useful for assessing how well the model distinguishes closely packed or irregularly shaped deals.

\begin{figure*}[h!]
    \centering
    \begin{tabular}{cc}
    \includegraphics[width=0.5\linewidth]{figures/deal_detection/map50_all_yolo.png} &   \includegraphics[width=0.5\linewidth]{figures/deal_detection/map50_95_all_yolo.png} \\
    (a) Results for mAP50 & (b) Results for mAP50-95 \\[2pt]
    \end{tabular}
    \caption{Performance comparison of YOLOv8 and YOLOv11 models on the deal detection validation set.}
    \label{fig:all_yolo_results}
\end{figure*}

The first set of experiments focused on comparing different YOLO variants trained on the dataset. Each model was trained for 20 epochs without additional data augmentation. The results, shown in Figure \ref{fig:all_yolo_results} (a), present the mAP50 (mask) scores for YOLOv8 and YOLOv11 in their small, medium, and large configurations. The findings indicate that the smallest version of YOLOv8 achieved the highest validation mAP50, outperforming larger models in this setting. A second analysis, visualized in Figure \ref{fig:all_yolo_results} (b), examines mAP50-95 scores for the same models, reinforcing the observation that a lightweight YOLO model is sufficient for this particular task. The lower performance of larger models suggests that the dataset’s structure, with relatively uniform layouts, does not require highly complex architectures to achieve strong results.

\begin{figure*}[h!]
    \centering
    \begin{tabular}{cc}
    \includegraphics[width=0.5\linewidth]{figures/deal_detection/map50_yolo8.png} &   \includegraphics[width=0.5\linewidth]{figures/deal_detection/map50_95_yolo8.png} \\
    (a) Results for mAP50 & (b) Results for mAP50-95 \\[2pt]
    \end{tabular}
    \caption{Performance comparison of YOLOv8s models with and without data augmentation on the deal detection validation set.}
    \label{fig:yolo8_results}
\end{figure*}

Following this baseline evaluation, additional experiments investigated the impact of data augmentation on segmentation performance. Various augmentation strategies were applied, including translation, scaling, horizontal flipping, and color transformations. The goal was to determine whether these transformations improved the model’s ability to generalize across different leaflet layouts. Each augmented model was trained for ten additional epochs, and the results are displayed in Figure \ref{fig:yolo8_results}. While certain augmentation techniques led to slight improvements in segmentation accuracy, particularly for models trained with random translations, the overall best-performing model remained the one trained without any augmentation. This outcome suggests that the dataset already contains sufficient variation to ensure robust generalization without the need for extensive artificial augmentations.

The final model selection was based on the combination of high segmentation accuracy and computational efficiency. Given the results, \emph{YOLOv8s} was chosen as the preferred model, as it consistently outperformed more complex variants while maintaining a lower computational cost. The model’s ability to detect deals with high precision makes it a strong candidate for further applications. And this is particularly important because everything else builds on it.

\section{Information Extraction of Supermarket Deals}
\label{sec:information_extraction}
Subsequently to the detection and separation of supermarket deals, the overarching goal of using these in applications requires the extraction of the various useful information from these deals. 
The task of extracting meaningful and structured information from supermarket deal images is a complex and multi-faceted problem, requiring the integration of various modalities, overcoming challenges posed by noisy data, and leveraging deep learning techniques. This section provides a detailed exploration of methodologies, as well as the theoretical and practical framework needed to perform such extraction.

\subsection{Approaches to Information Extraction}
Formally, the task of information extraction from a deal image can be defined as the identification of specific entities, including the product name, brand, original price, deal price, and unit. We define the following set of output labels:
\begin{equation}
\mathcal{Y} = \{y_{i}\}_{i=1}^{n},
\end{equation}
where $y_{i}$ represents the $i$-th entity in the deal image. In the following, the entities are defined as:
\begin{itemize}
    \item \code{$y_{\text{product\_name}}$}: The name of the product.
    \item \code{$y_{\text{brand}}$}: The brand of the product.
    \item \code{$y_{\text{original\_price}}$}: The original price of the product.
    \item \code{$y_{\text{deal\_price}}$}: The deal price of the product.
    \item \code{$y_{\text{unit}}$}: The unit of the product (e.g., weight, volume).
\end{itemize}

To extract these entities, one aims to learn a function $f(\cdot)$ that maps an input image $I$ to the desired outputs:
\begin{equation}
    f: I \to \mathcal{Y},
\end{equation}
where $ I $ is the deal image, and $ \mathcal{Y} $ represents the extracted entities.

The choice of the function $ f(\cdot) $ is an architectural decision that depends on the specific requirements of the task, the nature of the data, the available resources, and the desired performance metrics.

Among the various existing architectural approaches to learn $ f(\cdot) $ for information extraction, several methodologies have been developed to address the inherent complexities in supermarket deal images. These methods can be categorized into classical approaches, multi-stage traditional models, hybrid OCR-LLM systems, and end-to-end deep learning frameworks. Each paradigm offers distinct advantages, and their applicability depends on computational resources, dataset characteristics, and performance requirements.

\subsubsection{Traditional and Hybrid Approaches for Supermarket Leaflet Information Extraction}  
Information extraction from supermarket leaflets has traditionally been approached through a pipeline of sequential sub-tasks, each addressed by specialized systems. These methods, while extensively studied and methodologically diverse, are typically divided into three main stages: text detection, text recognition, and key-value pair extraction. Although effective in certain scenarios, these approaches suffer from inherent limitations, including computational overhead, error propagation, and suboptimal performance on edge cases. Recent advancements in large language models (LLMs) have led to the development of hybrid OCR-LLM architectures, which combine the precision of optical character recognition (OCR) with the contextual reasoning capabilities of LLMs. Below, we formalize these approaches and discuss their strengths and limitations.

Traditional architectures decompose the problem into independent sub-tasks, each handled by specialized models. While this modularity allows for targeted optimization, it introduces complexity and inefficiencies due to the lack of end-to-end training and error propagation across stages.

The first stage, \emph{text detection}, involves identifying regions in the image $ I \in \mathbb{R}^{H \times W \times 3} $ that contain textual information. This is typically achieved using rule-based systems or neural networks designed for small object detection. The output is a set of $ M $ bounding boxes $ \{b_i\}_{i=1}^{M} $, where each $ b_i = (x_{\text{min}}, y_{\text{min}}, x_{\text{max}}, y_{\text{max}}) $ defines the coordinates of a text region.

Each detected bounding box $ b_i $ is cropped and processed by a \emph{text recognition} model to convert the visual representation into a textual one. This task is challenging due to decoding errors caused by noisy backgrounds, stylized fonts, and ambiguous characters. Given the set of bounding boxes $ \{b_i\}_{i=1}^{M} $, the goal is to predict the corresponding textual entities $ \{t_i\}_{i=1}^{M} $ as the maximum likelihood sequence.

To use the recognized texts for information extraction, the final stage involves parsing the textual entities into structured outputs. This is typically achieved through rule-based systems or sequence-to-sequence models that generate key-value pairs from the recognized text. The output is a set of structured entities $ \{y_i\}_{i=1}^{N} $, where each $ y_i $ represents a specific entity (e.g., product name, price, unit).

The rise of \emph{large language models }(LLMs) has enabled the development of hybrid OCR-LLM architectures, which combine the strengths of OCR tools with the linguistic and reasoning capabilities of LLMs. These approaches retain the traditional OCR pipeline for converting visual text into machine-readable form but leverage LLMs to enhance robustness through contextual reasoning and structured output generation.

In a hybrid OCR-LLM system, the OCR stage processes the input image $ I $ to produce a set of unstructured text fragments $ T = \{t_i\}_{i=1}^{M} $ with associated spatial coordinates. These fragments are serialized into a prompt and fed into a pretrained LLM, which performs contextual parsing and generates some linguistic output that, in most cases, has to be somewhat postprocessed to extract the desired information.

Hybrid OCR-LLM models offer several advantages over traditional pipelines. By integrating contextual reasoning, they reduce error propagation and improve robustness to noisy or ambiguous inputs. Additionally, the use of pretrained LLMs eliminates the need for task-specific training, simplifying system development. However, these models still rely on the accuracy of the OCR stage, and misrecognitions (e.g., "O" vs. "0") can propagate to the LLM. Furthermore, the computational cost of LLM inference can be prohibitive for large-scale applications.

\subsubsection{End-to-End Models}  
End-to-End (E2E) approaches unify document understanding by learning a direct mapping $ F: I \to \mathcal{D} $ and thus jointly optimize visual feature extraction and semantic parsing. We formalize two dominant paradigms: vision-encoder-decoder architectures and large vision-language models (LVLMs), which share foundational principles but diverge in scale and architectural complexity.

Foundationally, all E2E models decompose into three components: a vision encoder, a representation projection, and a decoder. The vision encoder extracts hierarchical visual features $ \mathbf{Z} = \text{Enc}(I) $, which are then aligned with textual semantics through a representation projection. The decoder generates structured output $ \mathbf{Y} $ autoregressively. \figref{fig:e2e_arch} illustrates the unified architecture.

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.5\linewidth]{figures/vlm_arch.png}
    \caption{End-to-End Vision-Encoder-Decoder Architecture.}
    \label{fig:e2e_arch}
\end{figure}

\indent{\textbf{Donut \cite{kim2022}}.} This model employs a compact transformer-based encoder-decoder. Let $ I $ be partitioned into patches, each linearly projected to an encoded input $X$.  The used vision encoder $E$ computes an encoded representation mainly through:
\begin{equation}
Z = E(I) = \text{FFN}(\text{MSA}(X) + X) + \text{MSA}(X)
\end{equation}  
where $\text{MSA}$ denotes multi-head self-attention and $\text{FFN}$ a feed-forward network. The hierarchical structure captures local textural patterns as well as global layout cues.

Subsequently, a transformer decoder generates tokens $ Y = (y_1, \dots, y_T) $ using cross-attention over $ Z $:  
\begin{equation}
P(y_t | y_{<t}, Z) = \text{Softmax}(\text{FFN}(\text{MSA}(Y) + Z)),
\end{equation}  
where $ Y $ is the token sequence and $ Z $ the encoded image features. The model is trained end-to-end with a maximum likelihood objective \cite{vaswani2017}.

\indent{\textbf{Large Vision-Language Models (LVLMs)}.} In the realm of upscaling language models into LLMs, computer vision utilizes this trend to process multimodal data. LVLMs like Llama 3.2 Vision \cite{touvron2023} scale this framework by integrating pretrained vision encoders with large language models (LLMs). These architectures utilize \emph{cross-modal projection}, which aligns visual features with the LLM's embedding space, and \emph{multimodal decoding}, which attends to both visual tokens and textual prompts. The LVLM's strength lies in zero-shot reasoning and leveraging pretrained knowledge for structured output generation.

While smaller models like Donut offer efficiency and direct output generation, LVLMs provide zero-shot reasoning and open-world capabilities at increased computational cost.

\subsection{Dataset Creation}

\begin{table}[h!]
    \centering
    \caption{Leaflet-IE Dataset}
    \label{tab:ie_dataset}
    \begin{tabular}{lc}
    \toprule
    Entity           & Sample Size \\
    \midrule
    \code{image\_id}        & 372  \\
    \code{brand}            & 357  \\
    \code{product\_name}    & 370  \\
    \code{original\_price}  & 286  \\
    \code{deal\_price}      & 372  \\
    \code{weight}           & 369  \\
    \bottomrule
    \end{tabular}
    \end{table}

Since the availability of German supermarket leaflet data is literally not findable, a custom dataset was created by manually annotating a collection of supermarket deal images. The dataset includes images from the most common supermarket chains, each with unique layout, fonts, and colors to, in general, ensure the highest diversity and robustness possible. Each image is annotated with a corresponding label file that includes the desired entities as well as the image identificator. The dataset is split into a training (\~{} 80\%) and a validation set (\~{} 20\%) for each model to be trained and evaluated on. \tabref{tab:ie_dataset} provides an overview of the dataset with the sample size for each entity and \figref{fig:ie_dataset_samples} shows some sample images with their annotations.

Even though the Leaflet-IE dataset is relatively small, the reader will be able to see that the dataset is sufficient to evaluate the performance of different IE appraoches as well as to train an competitive end-to-end model. However, as the reader may notice, the dataset is solely focused on single product deals at this point and explicitly does not include

\begin{figure*}[h!]
\centering
\includegraphics[width=0.5\linewidth]{figures/ie_samples.png}
\caption{Sample images with annotations from the custom dataset.}
\label{fig:ie_dataset_samples}
\end{figure*}

\subsection{Experiments}
